---
title: 'Lab 03: Command Line Basics and Data Importing'
author: "Cindy Lo"
date: "9/12/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Basic Bash shell commands
```{r, eval = FALSE}
mkdir("~/Desktop/stat133/stat133-lab03")
cd ~/Desktop/stat133/stat133-lab03
curl -O http://textfiles.com/food/bread.txt
ls
curl -O http://textfiles.com/food/btaco.txt
curl -O http://textfiles.com/food/1st_aid.txt
curl -O http://textfiles.com/food/beesherb.txt
curl -O http://textfiles.com/food/bakebred.txt
curl -O http://archive.ics.uci.edu/ml/machine-learning-databases/forest-fires/forestfires.csv
curl -O http://web.pdx.edu/~gerbing/data/cars.csv
curl -O http://web.pdx.edu/~gerbing/data/color.csv
curl -O http://web.pdx.edu/~gerbing/data/snow.csv
curl -O http://web.pdx.edu/~gerbing/data/mid1.csv
curl -O http://web.pdx.edu/~gerbing/data/mid2.csv
curl -O http://web.pdx.edu/~gerbing/data/minutes1.csv
curl -O http://web.pdx.edu/~gerbing/data/minutes2.csv

ls -l #to list the contents in your current directory in long format
man ls #Look at the `man` documentation of `ls`
ls -r #list the contents in reverse order
ls -l -t
ls *.txt
ls -r *.csv
ls ????.csv
ls *[2]*.csv
ls *[[:digit:]]*.csv #or `ls *[0-9]*.csv`
mkdir data
cd data
mkdir txt-files
mkdir csv-files
mv ../bread.txt  txt-files
mv ../*.txt  txt-files
mv ../*.csv csv-files
cd ../..
cp data/txt-files/bread.txt copies
cp data/txt-files/*.txt copies
cp data/csv-files/*.csv copies
cd copies
mv bread.txt bread-recipe.txt
mv Fisher.csv iris.csv #there's no Fisher.csv in our directory, so I wrote what I would write down if it did exist
mv btaco.txt breakfast-taco.txt
cd ..
mv copies copy-files
rm copy-files/*.csv
rm -R copy-files
ls -r data/txt-files
```

## Abalone Data Set
```{r}
url <- "http://archive.ics.uci.edu/ml/machine-learning-databases/abalone/abalone.data"
abalone <- read.table(url, sep = ",")
```
What's the character delimiter?

- The character delimiter is a comma (,).

Is there a row for column names?

- There is no row for column names.

Are there any missing values? If so, how are they codified?

- Missing values were removed because the majority had the predicted value missing. The ranges of the continuous values have been scaled for use with an ANN (by dividing by 200).

What is the data type of each column?

- The first column is nominal. Columns 2 to 8 have continuous data type, and the last column has an integer data type.

### Getting a Local Copy of the Data
```{r}
#origin <- 'http://archive.ics.uci.edu/ml/machine-learning-databases/abalone/abalone.data'
#destination <- 'abalone.data'
#download.file(origin, destination)
ls()
```

## Some Bash Commands
```{r, eval = FALSE}
file abalone.data
```
`abalone.data` is a ASCII text

```{r, eval = FALSE}
wc abalone.data
```
`abalone.data` has

  1. newline count of 4177
  2. word count of 4177
  3. byte count of 191873
```{r, eval = FALSE}
man wc
wc -l abalone.data
head abalone.data
man head
head -n 5 abalone.data
tail abalone.data
man tail
tail -n 3 abalone.data
less abalone.data
q
```

## Basic Importing
```{r}
abalone <- read.table("abalone.data", sep = ",")
head(abalone)
tail(abalone)
str(abalone, vec.len = 1)
```

### Detailed information about the columns
```{r}
column_names <- c("Sex", "Length", "Diameter", "Height", "Whole weight", "Shucked weight", "Viscera weight", "Shell weight", "Rings")

column_types <- c("character", "numeric", "real", "real", "real", "real", "real", "real", "integer")

abalone <- read.table("abalone.data", sep = ",", col.names = column_names, colClasses = column_types)
str(abalone)

abalone2 <- read.csv("abalone.data", sep = ",", col.names = column_names, colClasses = column_types)
str(abalone2)

abalone10 <- head(abalone, 10)
str(abalone10)

abalone20 <- head(abalone, skip = 10, 10)
str(abalone20)

summary(abalone)
apply(abalone, 2, sd)

